---
title: "Homework 1"
author: Vera Xu
output: github_document
---

This is my solution to HW1.

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Problem 1

Create a data frame with the specified elements. 

```{r create_df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high"))
  )
```

Take the mean of each variable in my data frame.

```{r mean_pull_var}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

I can take the mean of numbers and logical but not character or factor. 

```{r try_numeric}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

```{r}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
```


## Problem 2

```{r}
data("penguins", package = "palmerpenguins")
```

#### Description of the penguins dataset
* The penguins dataset contains the following variables: `r ls(penguins)`.
  * The species variable contains `r unique(pull(penguins, species))`
  * The island variable contains `r unique(pull(penguins, island))`
* The penguins dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.
* The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

#### Scatterplot showing the distribution of bill length and flipper length by species
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species))+ geom_point()
```

```{r}
ggsave("p8105_hw1_q2.png")
```

